---
title: "ESM 244 Lab 1"
author: "Taylor Medina"
date: "1/5/2022"
output: html_document
---

```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Reading in data
```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Part 1: Wrangling and ggplot review

**Example 1:** Find count of observations by `legal_status` and wrangle data

```{r}
# Method 1
sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())

# Method 2
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>%  
  slice_max(tree_count, n = 5) %>% 
  arrange(-tree_count)
```

Make a graph of top 5 from above:
```{r}
ggplot(data = top_5_status, aes(
  x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col(fill = 'darkgreen') +
  labs(x = 'legal status',
       y = 'tree counts') +
  coord_flip() +
  theme_minimal()
```

**Example 2:** Only going to keep observations where legal status is 'Permitted Site' and caretaker is 'MTA'

```{r}
permitted_data_sf <- sf_trees %>% 
  filter(legal_status == 'Permitted Site', caretaker == 'MTA')
```

**Example 3:** Only keep blackwood acacia trees and only keep columns `legal_status`, `date`, `latitude`, and `longitude`.

```{r}
blackwood_acacia_df <- sf_trees %>%
  filter(str_detect(species, 'Blackwood Acacia')) %>%
  select(legal_status, date, lat = latitude, lon = longitude)
```

### Make a graph of locations
```{r}
ggplot(data = blackwood_acacia_df, aes(x = lon, y = lat)) +
  geom_point(color = 'darkgreen')
```

