---
title: "ESM 244 Lab 1"
author: "Taylor Medina"
date: "1/5/2022"
output: html_document
---

```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Reading in data
```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Part 1: Wrangling and ggplot review

**Example 1:** Find count of observations by `legal_status` and wrangle data

```{r}
# Method 1
sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())

# Method 2
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>%  
  slice_max(tree_count, n = 5) %>% 
  arrange(-tree_count)
```

Make a graph of top 5 from above:
```{r}
ggplot(data = top_5_status, aes(
  x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col(fill = 'darkgreen') +
  labs(x = 'legal status',
       y = 'tree counts') +
  coord_flip() +
  theme_minimal()
```

**Example 2:** Only going to keep observations where legal status is 'Permitted Site' and caretaker is 'MTA'

```{r}
permitted_data_sf <- sf_trees %>% 
  filter(legal_status == 'Permitted Site', caretaker == 'MTA')
```

**Example 3:** Only keep blackwood acacia trees and only keep columns `legal_status`, `date`, `latitude`, and `longitude`.

```{r}
blackwood_acacia_df <- sf_trees %>%
  filter(str_detect(species, 'Blackwood Acacia')) %>%
  select(legal_status, date, lat = latitude, lon = longitude)
```

Make a graph of locations
```{r}
ggplot(data = blackwood_acacia_df, aes(x = lon, y = lat)) +
  geom_point(color = 'darkgreen')
```

**Example 4:** `tidyr::separate()`
```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c('spp_scientific',
                             'spp_common'),
           sep = ' :: ')
```

**Example 5:** `tidyr::unite()`
```{r}
ex_5 <- sf_trees %>% 
  unite('id_status', tree_id, legal_status, sep = ' ')
```

### Part 2: Make some maps

Convert lat/long data to spatial points, `sf::st_as_sf()`
```{r}
blackwood_acacia_sf <- blackwood_acacia_df %>% 
  drop_na(lat, lon) %>% 
  st_as_sf(coords = c('lon', 'lat'))

# Tell R coordinate reference system
st_crs(blackwood_acacia_sf) <- 4326

ggplot(data = blackwood_acacia_sf) +
  geom_sf(color = 'dark green') +
  theme_minimal()
``` 

Read in SF shapefile and add to map
```{r}
sf_map <- read_sf(here('data', 'sf_map', 'tl_2017_06075_roads.shp'))

sf_map_transform <- st_transform(sf_map, 4326)

ggplot(data = sf_map_transform) +
  geom_sf()
```

Combine maps
```{r}
ggplot() +
  geom_sf(data = sf_map, size = 0.1, color = 'darkgrey') +
  geom_sf(data = blackwood_acacia_sf, size = 0.5, color = 'red') +
  theme_void() +
  labs(title = 'Blackwood Acacia Trees in SF')
```

### Interactive Map
```{r}
tmap_mode('view')

tm_shape(blackwood_acacia_sf) +
  tm_dots()
```

